A sample article title

### 2019-05-15 bmc_article.tex
...

\subsection*{Doublet detection}

@@

\subsection*{Filtering: excluded more cells is not necessarily good}

@@

On the basis of these observations, we defined three sets of criteria as lenient, default and stringent filtering (see methods), and investigated them further.

@@

>>>>

\subsection*{Filtering}

\subsubsection*{Doublet detection}

@@

\subsubsection*{Excluded more cells is not necessarily good}

@@

On the basis of these observations, we defined three sets of criteria as lenient, default and stringent filtering (see methods), and investigated them further. [...]

\subsubsection*{Filtering features by type}

Mitochondrial reads have been associated with cell degradation, and it has been observed that much of scRNAseq clustering is determined by ribosomal genes [REF]. We therefore considered whether excluding one category of features or the other, or using only protein-coding genes, had impact on the ability to distinguish subpopulations.

@@

We next investigated the impact of different normalization strategies. In addition to log-normalization, the standard Seurat clustering pipeline scales features so that the PCA is not dominated by a small subset of the features, and we therefore included versions of the different normalization, with or without a subsequent scaling (except sctransform's variance-stabilizing transformation, given that, conceptually, it involves something analogous to scaling). Finally, Seurat's `ScaleData` includes the option to regress out the effect of certain covariates, and we therefore tested its use with the proportion of mitochondrial counts and the number of detected features.

An important motivation of sctransform was the observation that, even after normalization, the first principal components of various datasets tended to correlate with library size, suggesting an imperfect correlation \citep{hafemeister_sctransform_2019}. We therefore first assessed to what extent the first principal component still retained a correlation with the library size and number of detected features (Supplementary XX). In the 10X datasets, most methods successfully decreased the correlation with covariates, including the mere scaling of non-normalized data. This was not the case in the smartseq datasets, which required the additional regressing out of covariates.

(NOTE: a correlation of PC1 with library size might be caused by the fact that subpopulations have a different average library size. I'll therefore repeat this evaluation with the variance explained by libSize in a PC1~subpopulation+libSize model.)

We next investigated the impact of normalization on the ability to distinguish subpopulations, which was much greater than that of filtering (Supplementary Figure XX). For many clusters, scaling tended to reduce both the average silhouette width (XXX) and the proportion of variance in the top principal components explained by subpopulations (Figure 3 and Supplementary Figure XXX), and was associated with a poorer adjusted Rand index (ARI) of the resulting clustering (Figure 3). Regressing out covariates systematically gave poorer performance on all metrics. sctransform produced the highest cluster average silhouette width (Supplementary Figure XXX) as well as ARI (Figure 3), but not always the highest mean proportion of variance in the top components explained by subpopulations (Figure 3 and Supplementary Figure XXX). Interestingly, clusters with the lowest average silhouette width (and which are therefore more difficult to classify) tended to be less well distinguished with sctransform, and instead showed a good difference in non-normalized and/or non-scaled data, suggesting that some procedures emphasize differences more useful to track some subpopulations than others, and hence the possibility that further developments in methods could improve this step.

\subsection*{Feature selection}

A standard clustering pipeline typically involves a step of selecting highly-variable genes, which is complicated by the digital nature and the mean-variance relationship of RNAseq. Seurat's earlier approaches involved the use of dispersion estimates standardized for the expression levels, while more recent versions rely of a different measure of variance, again standardized. Of note, while adjusting for the mean-variance relationship removes much of the bias towards highly-expressed genes, it is nevertheless plausible that this bias reflects biological relevance and hence is actually helpful in classifying cell types. \citep{townes_glmpca_2019}	instead suggested to use deviance, while sctransform provides its own ordering the genes by transformed variance. We therefore compared these methods.

Reasoning that a selection method should ideally select genes whose variability is more between subpopulations than within, we first assessed whether each selection method selected genes with a high proportion of variance or deviance explained by (real) subpopulations (the latter estimated through edgeR). Estimates of variance explained by subpopulations based on a standard Seurat normalization or on sctransform data were highly correlated (Supplementary Figure XXX A), and there was also a good agreement with deviance explained, although lowly-expressed genes in particular could have a high deviance explained without having much of their variance explained by subpopulations (Supplementary Figure XXX B-D). We therefore compared, using both of these estimates, the proportion of the cumulative variance/deviance explained by the top X genes that could be retrieved by each gene ranking method (Supplementary Figures XX and XX). We focused on the first 1000 genes to highlight differences between methods, although the differences tended to decrease with a higher number of genes selected. A first observation was that standardized measures of variability were systematically worse than their non-standardized  counterparts in selecting genes with a high proportion of variance explained by subpopulations, while instead proving most of the time superior in selecting genes with a high _deviance_ explained by subpopulations (Figure 4 and Supplementary Figure XX). Deviance proved the method of choice to maximize the former (with mere expression level proving surprisingly good), but did not perform so well to select genes with a high deviance explained.

We next investigated whether a selection based on these methods impacted the proportion of variance explained by subpopulations in the first principal components (Supplementary Figure XX). To validate the previous assay, included a selection based on the genes whose proportion variance or deviance explained was highest. Interestingly, while these selections were on average the top-ranking methods, they were not systematically so for all dataset.

These main observations were reflected in the ARI of the resulting clustering (Figure 4B): non-standardized measures of variability, including mere expression level, tended to outperform more complex metrics. In general, we found Seurat's unstandardized estimate of variance to provide the best results across all normalization methods, followed by deviance.

We next investigated the impact of the number of features selected.

\subsection*{Dimensionality reduction}
